### 2016-11-18 capability-machine.tex
Added lemmas we expect to hold
changed the awkward example slightly

%World operations
\newcommand{\future}{\mathbin{\sqsupseteq}}
\newcommand{\futurewk}{\mathbin{\sqsupseteq}_{\var{wk}}}
\newcommand{\futurestr}{\mathbin{\sqsupseteq}_{\var{str}}}
\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}

@@

\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
\[
\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\memheap&\in& \Heaps \\
\memreg &\in& \Regs \\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
n       &\in& \ints\\
\end{array}
\]
Further definitions:
\[
\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& 
             \jmp{\lv} \mid 
             \jnz{\lv}{\rv} \mid
             \move{\lv}{\rv} \mid 
             \load{\lv}{\hv} \mid 
             \store{\hv}{\rv} \mid  \\
       &   & \plus{\lv}{\rv}{\rv} \mid 
             \lea{\lv}{\rv} \mid 
             \restricttwo{\lv}{\rv} \mid 
             \subseg{\lv}{\rv}{\rv} \mid  \\
       &   & \isptr{\lv}{\rv} \mid 
             \getp{\lv}{\lv} \mid 
             \getl{\lv}{\lv} \mid 
             \getb{\lv}{\lv} \mid
             \gete{\lv}{\lv} \mid
             \geta{\lv}{\lv} \mid \\
       &   & \fail \mid
             \halt \mid 
\end{array}
\]

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x:=0;
                 f();
                 x:=1;
                 f();
                 assert(x==1)
\end{verbatim}
\begin{lemma}
  \newcommand{\adv}{\var{adv}}
  \newcommand{\flag}{\var{flag}}
  For all $n \in \nats$ 
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus h_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(hs_g) = [\start_g,\addrend_g] \\
  &\dom(hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\adv}]{n}{\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}}
\end{align*}
and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$, 
then
\[
  h'(\flag) = 0
\]
\end{lemma}

\subsection{Ticket Dispenser}
\label{sec:tick-disp}
\newcommand{\hsfoot}{\hs_\var{footprint}}
\newcommand{\hsframe}{\hs_\var{frame}}
\newcommand{\size}{\var{size}}
\newcommand{\rio}{r_{io}}
\newcommand{\adv}{\codelabel{adv}}
\newcommand{\advb}{\var{adv_{base}}}
\newcommand{\adve}{\var{adv_{end}}}
\newcommand{\initb}{\var{init}_{base}}
\newcommand{\inite}{\var{init}_{end}}
\newcommand{\mrlen}{5cm}
\newcommand{\retm}{\var{ret}_{\malloc}}
\newcommand{\reta}{\var{ret}_{\adv}}
\newcommand{\base}{\var{base}}
\newcommand{\eend}{\var{end}}
\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
 \left . \vphantom{\begin{array}{l}
 \ifthenelse{\equal{#1}{1}}{3\\}{
    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
      3\\3\\3\\3\\3\\3\\3\\ %7
  }}}}}}
  \end{array}} \right \}}}
}
\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
\lau{Has not been adjusted since local capabilities were added.}

@@

Standard regions for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}
Revoking all temporary regions:
\begin{align*}
  \revokeTemp{} & : \Worlds \fun \Worlds \\
  \revokeTemp{W} & \defeq \lambda r \ldotp 
                   \begin{cases}
                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
                     W(r)                & \text{otherwise}
                   \end{cases}
\end{align*}
Standard regions for when write local is not allowed:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
    \begin{cases}
      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
      \bot & \text{otherwise}
\end{cases}
\\ \\
  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

>>>>

%World operations
\newcommand{\future}{\mathbin{\sqsupseteq}}
\newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
\newcommand{\futurestr}{\mathbin{\sqsupseteq}^{\var{priv}}}
\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}

@@

\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
\[
\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\memheap&\in& \Heaps \\
\memreg &\in& \Regs \\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
n       &\in& \ints\\
\end{array}
\]
Further definitions:
\[
\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& 
             \jmp{\lv} \mid 
             \jnz{\lv}{\rv} \mid
             \move{\lv}{\rv} \mid 
             \load{\lv}{\hv} \mid 
             \store{\hv}{\rv} \mid  \\
       &   & \plus{\lv}{\rv}{\rv} \mid 
             \lea{\lv}{\rv} \mid 
             \restricttwo{\lv}{\rv} \mid 
             \subseg{\lv}{\rv}{\rv} \mid  \\
       &   & \isptr{\lv}{\rv} \mid 
             \getp{\lv}{\lv} \mid 
             \getl{\lv}{\lv} \mid 
             \getb{\lv}{\lv} \mid
             \gete{\lv}{\lv} \mid
             \geta{\lv}{\lv} \mid \\
       &   & \fail \mid
             \halt \mid 
\end{array}
\]
Further define $\reg_0 \in \Regs$ such that
\[
  \forall r \in \RegName \ldotp \reg_0(r) = 0
\]

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x:=0;
                 f();
                 x:=1;
                 f();
                 assert(x==1)
\end{verbatim}
\begin{lemma}\forcenewline
  \newcommand{\adv}{\var{adv}}
  \newcommand{\flag}{\var{flag}}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus h_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(hs_g) = [\start_g,\addrend_g] \\
  &\dom(hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}]}
\end{align*}
and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$ as well as an entry capability for $\malloc$, 
then
\[
  h'(\flag) = 0
\]
\end{lemma}

Standard regions for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}
Revoking all temporary regions:
\begin{align*}
  \revokeTemp{} & : \Worlds \fun \Worlds \\
  \revokeTemp{W} & \defeq \lambda r \ldotp 
                   \begin{cases}
                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
                     W(r)                & \text{otherwise}
                   \end{cases}
\end{align*}
\newcommand{\wrev}[1]{#1_{\var{rev}}}
Let
\[
\revokeTemp(W) = \wrev{W}
\]

Standard regions for when write local is not allowed:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
    \begin{cases}
      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
      \bot & \text{otherwise}
\end{cases}
\\ \\
  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

\subsection{Lemmas}
We expect the following lemmas to hold true:
\subsubsection{Heap Satisfaction}
\begin{lemma}[Downwards closure]
  \begin{align*}
    &\forall \hs, n' leq n, W \ldotp \\
    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
  \end{align*}
\end{lemma}

\begin{lemma}[Public monotonicity]
  \begin{align*}
    & \forall \hs, n, W' \futurewk W \ldotp \\
    & \quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n}{W'}
  \end{align*}
\end{lemma}

\begin{lemma}[Private monotonicity-like property]
  \begin{align*}
    & \forall \hs, n, W, W' \ldotp \\
    & \quad  \heapSat[\hs]{n}{W} \land W' \futurewk \wrev{W} \futurestr W \\
    & \qquad \Rightarrow \exists \hs', \hs_f \ldotp \hs = \hs' \uplus \hs_f \land \heapSat[\hs']{n}{W'}
  \end{align*}
\end{lemma}

\subsubsection{Future world}
\begin{lemma}
  For all $W$, $W'$, and $W''$
  \begin{itemize}
  \item $W' \futurewk W \Rightarrow W' \futurestr W$
  \item $\wrev{W} \futurestr W$
  \item $W \futurestr \wrev{W} \Rightarrow W \futurewk \wrev{W}$
  \end{itemize}
  Combined with transitivity of $\futurestr$, the above gives us:
  \begin{itemize}
  \item $W'' \futurestr W' \land W' \futurewk W \Rightarrow W'' \futurestr W$
  \item $W'' \futurewk W' \land W' \futurestr W \Rightarrow W'' \futurestr W$
  \end{itemize}
\end{lemma}

\subsubsection{Value relation}
\begin{lemma}[Global capabilities monotone wrt $\futurestr$]
  \begin{align*}
    & \forall n, \perm, \start, \addrend, \addr, W \ldotp \\
    & \quad  \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W) \\
    & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(\wrev{W}) 
  \end{align*}
\end{lemma}

\section{Old TD example}
\subsection{Ticket Dispenser}
\label{sec:tick-disp}
\newcommand{\hsfoot}{\hs_\var{footprint}}
\newcommand{\hsframe}{\hs_\var{frame}}
\newcommand{\size}{\var{size}}
\newcommand{\rio}{r_{io}}
\newcommand{\adv}{\codelabel{adv}}
\newcommand{\advb}{\var{adv_{base}}}
\newcommand{\adve}{\var{adv_{end}}}
\newcommand{\initb}{\var{init}_{base}}
\newcommand{\inite}{\var{init}_{end}}
\newcommand{\mrlen}{5cm}
\newcommand{\retm}{\var{ret}_{\malloc}}
\newcommand{\reta}{\var{ret}_{\adv}}
\newcommand{\base}{\var{base}}
\newcommand{\eend}{\var{end}}
\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
 \left . \vphantom{\begin{array}{l}
 \ifthenelse{\equal{#1}{1}}{3\\}{
    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
      3\\3\\3\\3\\3\\3\\3\\ %7
  }}}}}}
  \end{array}} \right \}}}
}
\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
\lau{Has not been adjusted since local capabilities were added.}

@@
